---
title: "Riddler Express Solution: Holiday Party Voting"
author: "Quoc Tran"
date: "12/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [Riddler Express](https://fivethirtyeight.com/features/can-you-find-a-matching-pair-of-socks/)

#### By Zach Wissner-Gross

From Mark Hannan comes a holiday party stumper:

You’re new at your job, and your office is voting on a theme for its holiday party. It’s fallen on you to record the percent of your coworkers (including yourself) who voted for each one. Well, since you’re in a hurry, you just write down everything in the percentage that comes before the decimal point. So for example, 35.0 percent, 35.17 percent and 35.92 percent would all be written simply as “35 percent.”

After the votes are tallied, you found that the winner received 73 percent of the vote (at least, that’s what you wrote down), while second place had 58 percent, and third place had 32 percent. Your first realization is that you work with a bunch of cheaters who voted more than once. But your second thought is that you might be able to use this information to figure out how many people work in your office. (As I said, you’re new, and this isn’t something you know off the top of your head.)

Based on these percentages, what’s the minimum number of people who could work in your office?

*Extra credit*: Your office could be filled with many possible numbers of people. Based on the percentages given in the problem, what’s the greatest number of people your office *can’t* have?

***

## My Solution

To find how many $x$ number of people work in the office, we just need to find all values of $x$ in this system of equations that result in whole numbers for $a$, $b$, and $c$:

$$\begin{cases} 0.32x \leq a \leq 0.33x \\ 0.58x \leq b \leq 0.59x \\ 0.73x \leq c \leq 0.74x \end{cases}$$

We just need to find the lowest number of people in the office, so we can just run through for all positive integers of $x$ from $1$ to $1000$ (anything greater than that is absurd; no office party can host that many people), with a resolution of $0.000001$ between each percentage point (that seems sufficiently small enough).

```{r}
for (x in 1:1000) {
  testA <- c()
  for (a in seq(0.33,0.34,0.000001)) {
    if ((x*a) %% 1 == 0) {
      testA <- c(testA,x*a)
    }
  }
  testB <- c()
  for (b in seq(0.58,0.59,0.000001)) {
    if ((x*b) %% 1 == 0) {
      testB <- c(testB,x*b)
    }
  }
  testC <- c()
  for (c in seq(0.73,0.74,0.000001)) {
    if ((x*c) %% 1 == 0) {
      testC <- c(testC,x*c)
    }
  }
  if (length(testA) != 0 && length(testB) != 0 && length(testC) != 0) {
    print(paste("There can be ", x, " people in the office.", sep=""))
    print("Votes for option A:")
    print(testA)
    print("Votes for option B:")
    print(testB)
    print("Votes for option C:")
    print(testC)
  }
}
```

