---
title: "Riddler Classic Solution: Can You Beat The GOAT Monty Hall Problem?"
author: "Quoc Tran"
date: "04/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### [Riddler Classic](https://fivethirtyeight.com/features/can-you-beat-the-goat-monty-hall-problem/)

#### By Zach Wissner-Gross

The [Monty Hall problem](https://www.scientificamerican.com/article/how-randomness-rules-our-world/) is a classic case of conditional probability. In the original problem, there are three doors, two of which have goats behind them, while the third has a prize. You pick one of the doors, and then Monty (who knows in advance which door has the prize) will always open another door, revealing a goat behind it. It’s then up to you to choose whether to stay with your initial guess or to switch to the remaining door. Your best bet is to switch doors, in which case you will win the prize two-thirds of the time.

Now suppose Monty changes the rules. First, he will randomly pick a number of goats to put behind the doors: zero, one, two or three, each with a 25 percent chance. After the number of goats is chosen, they are assigned to the doors at random, and each door has at most one goat. Any doors that don’t have a goat behind them have an identical prize behind them.

At this point, you choose a door. If Monty is able to open another door, revealing a goat, he will do so. But if no other doors have goats behind them, he will tell you that is the case.

It just so happens that when you play, Monty is able to open another door, revealing a goat behind it. Should you stay with your original selection or switch? And what are your chances of winning the prize?

***

## My Solution

It's obvious that I'm going to end up doing this with Monte-Carlo experiments anyways, so we can just start off with that.

```{r}

playGame <- function(strategy) {
  numGoats <- sample(0:3, 1)
  doorGoats <- c(rep.int(TRUE, numGoats), rep.int(FALSE, 3 - numGoats))
  doorGoats <- sample(doorGoats, 3)
  
  doorChoice <- sample(1:3, 1)
  
  if(TRUE %in% doorGoats[-doorChoice]) { # Goats still remain in unopened doors
    if (strategy == "switch") {
      if (numGoats == 1 ) {
        return(TRUE)
      } else if (numGoats == 2) {
        return(sample(c(TRUE, FALSE), 1))
      } else if (numGoats == 3) {
        return(FALSE)
      }
    } else if (strategy == "stay") {
      if (doorGoats[doorChoice] == FALSE) {
        return(TRUE)
      } else {
        return(FALSE)
      }
    }
  } else { # Either no goats or you've chosen the only goat
    return(NA)
  }
}

```


```{r}
switchGames <- c()
stayGames <- c()
for (i in 1:10000) {
  switchGames <-c(switchGames, playGame("switch"))
  stayGames <-c(stayGames, playGame("stay"))
}

mean(switchGames, na.rm = TRUE)
mean(stayGames, na.rm = TRUE)

```

